---
title: "Code<br><br>"
menu: main
output: 
  html_document:
    df_print: paged
    toc: true
    toc_depth: 2
    toc_float: true
    number_sections: false
    theme: united
    highlight: espresso
---

```{r, include=FALSE}
setwd("~/Desktop/College/Senior Year/Spring/Data analytics/Data Project/archive")

library(socviz)
library(lubridate)
library(geofacet)
library(ggthemes)
library(ggrepel)
library(ggridges)
library(plyr)
library(tidyverse)
library(plotly)
library(gapminder)
library(stargazer)
theme_set(theme_minimal())
```
# Loading the Data

```{r}
# Read 2015 Data
h15 <- read_csv("2015.csv")
h15 <- h15 %>%
  dplyr::mutate(Year = 2015) %>%
  dplyr::rename(H_rank=`Happiness Rank`, # Modify variable names
                H_score = `Happiness Score`,
                GDP=`Economy (GDP per Capita)`,
                Health=`Health (Life Expectancy)`,
                Trust=`Trust (Government Corruption)`,
                SE=`Standard Error`,
                dystopia_res = `Dystopia Residual`)


# Read 2016 Data
h16 <- read_csv("2016.csv")  
h16 <- h16 %>%
  dplyr::mutate(Year = 2016,
      `Standard Error` = (`Upper Confidence Interval`-`Lower Confidence Interval`)/3.92) %>%
              # SE = (upper limit – lower limit) / 3.92. 
              # This is for 95% CI
  dplyr::select(-c(`Upper Confidence Interval`,`Lower Confidence Interval`)) %>%
  dplyr::rename(H_rank=`Happiness Rank`, # Modify variable names
                H_score = `Happiness Score`,
                GDP=`Economy (GDP per Capita)`,
                Health=`Health (Life Expectancy)`,
                Trust=`Trust (Government Corruption)`,
                SE=`Standard Error`,
                dystopia_res = `Dystopia Residual`)



# Since we don't have a variable 'Region' starting from 2017, we will create it for 
# each year
h_regions <- dplyr::select(h16, Country, Region)



# Read 2017 Data
h17 <- read_csv("2017.csv")  
h17 <- h17 %>%
  dplyr::mutate(Year = 2017,
         `Standard Error` = (`Whisker.high`-`Whisker.low`)/3.92) %>%
  merge(h_regions,by="Country", all.x=T) %>%
  dplyr::select(-c(`Whisker.high`,`Whisker.low`)) %>%
    dplyr::rename(H_rank=`Happiness.Rank`, # Modify variable names
                H_score = Happiness.Score,
                GDP=Economy..GDP.per.Capita.,
                Health=Health..Life.Expectancy.,
                Trust=Trust..Government.Corruption.,
                SE=`Standard Error`,
                dystopia_res = Dystopia.Residual)


# Read 2018 Data
h18 <- read_csv("2018.csv")  
h18 <- h18 %>%
  dplyr::mutate(Year = 2018) %>%
  dplyr::rename(H_rank=`Overall rank`, # Modify variable names
                H_score = `Score`,
                GDP=`GDP per capita`,
                Country = `Country or region`,
                Health=`Healthy life expectancy`,
                Trust=`Perceptions of corruption`,
                Freedom = `Freedom to make life choices`,
                Family = `Social support`) %>%
  merge(h_regions,by="Country", all.x=T)


# Read 2019 Data
h19 <- read_csv("2019.csv")  
h19 <- h19 %>%
  dplyr::mutate(Year = 2019) %>%
  dplyr::rename(H_rank=`Overall rank`, # Modify variable names
                H_score = `Score`,
                GDP=`GDP per capita`,
                Country = `Country or region`,
                Health=`Healthy life expectancy`,
                Trust=`Perceptions of corruption`,
                Freedom = `Freedom to make life choices`,
                Family = `Social support`) %>%
  merge(h_regions,by="Country", all.x=T)

# Combine all data into all_dat
h_alldat <- tibble(rbind.fill(h15,h16,h17,h18,h19))

glimpse(h_alldat)
```

## Merging Happiness and Country Datasets

```{r}
# Merging the world dataset and happiness sets together
library(readr)
countries <- read_csv("~/Downloads/countries of the world.csv")

# All happiness data merged with country data. Country data is over 30 years
# so it is the same for each year.

ch_alldat <- merge(h_alldat, countries, by ="Country")
ch15 <- merge(h15, countries, by = "Country")
ch16 <- merge(h16, countries, by = "Country")
ch17 <- merge(h17, countries, by = "Country")
ch18 <- merge(h18, countries, by = "Country")
ch19 <- merge(h19, countries, by = "Country")
```

```{r}
library(janitor)

ch15 <- ch15 %>%
  clean_names()

ch_alldat <- ch_alldat %>%
  clean_names()
```

```{r}
head(ch_alldat)
```
# Top 10 Happiest Countries

```{r}
# Top 10 happiest countries 2015 - 2019 (Mean)
# 
# ( top_10 <- ch_alldat %>%
#   group_by(country) %>%
#   dplyr::summarise(mean_rank = mean(h_rank)) %>%
#   arrange(desc(mean_rank)) %>%
#   filter(mean_rank <= 10) )

```

# Visualizations

## Happiness Over Time

I want to show how each region's mean happiness score has trended over the 5 year period.
Need to group by region and year and find the mean of each.

```{r}

reg_mean <- ch_alldat %>%
  group_by(region_x, year) %>%
  dplyr::summarise(mean(h_score))


p <- ggplot(data = reg_mean, mapping = aes(x= `mean(h_score)`, y= region_x, color = year)) +
  geom_col() + theme_bw() + labs(x = "Year", y = "Happiness Score",
                                   color = "Region")
p

```

## Happiness and Phones  


```{r}
# ggplot(data = ch15, mapping = aes(y = h_score, x = phones_per_1000, color = region_x )) +
#   geom_point()

p <- ch_alldat %>%
  filter(year == 2015) %>%
  ggplot(aes(phones_per_1000, h_score, color = region_x, label = country)) +
  geom_point() + theme_bw() + labs(x = "Phones per 1000", y = "Happiness Score",
                                   color = "Region")

p1 <- ch_alldat %>%
  filter(year == 2019) %>%
  ggplot(aes(phones_per_1000, h_score, color = region_x, label = country)) +
  geom_point() + theme_bw() + labs(x = "Phones per 1000", y = "Happiness Score",
                                   color = "Region")

ggplotly(p)
ggplotly(p1)
```

## Happiness and Literacy

```{r}
p <- ch_alldat %>%
  filter(year == 2015) %>%
  ggplot(aes(literacy_percent, h_score, color = region_x, label = country)) +
  geom_point() + theme_bw() + labs(x = "Literacy Percent", y = "Happiness Score",
                                   color = "Region")
ggplotly(p)
```

## Happiness and GDP per Capita
```{r}
p <- ch_alldat %>%
  filter(year == 2015) %>%
  ggplot(aes(gdp_per_capita, h_score, color = region_x, label = country)) +
  geom_point() + theme_bw() + labs(x = "GDP per Capita", y = "Happiness Score",
                                   color = "Region")
ggplotly(p)
```

```{r}
p <- ch_alldat %>%
  filter(year == 2015) %>%
  ggplot(aes(dystopia_res, h_score, color = region_x, label = country)) +
  geom_point() + theme_bw() + labs(x = "Dystopia Residual", y = "Happiness Score",
                                   color = "Region")
ggplotly(p)
```
## Infant Mortality and Happiness

```{r}
p <- ch_alldat %>%
  filter(year == 2015) %>%
  ggplot(aes(infant_mortality_per_1000_births, h_score, color = region_x, label = country)) +
  geom_point() + theme_bw() + labs(x = "Infant Mortality (per 1000)", y = "Happiness Score",
                                   color = "Region")
ggplotly(p)
```


# Mean Happiness Score by Region

```{r}
regional_h <- ch_alldat %>%
  group_by(region_x) %>%
  dplyr::summarise(mean_score = mean(h_score)) %>%
  arrange(desc(mean_score))

reg_h_plot <- ggplot(regional_h, mapping = aes(x = mean_score, y = region_x, fill = region_x, 
                                               label = mean_score)) + geom_col()
reg_h_plot
```


# Modeling the Data

For creating my model, I wanted to take a deeper dive into the 3 variables I thought 
would be most influential. 

```{r results = "asis", echo = TRUE, eval=T}
model <- lm(h_score ~ gdp_per_capita + literacy_percent + infant_mortality_per_1000_births,
            data = ch_alldat)

stargazer(model, type = "html", omit = c("Constant"))

```

These 3 variables seem to have the highest correlation with happiness scores. GDP per capita, literacy percent, and infant mortality rates do make sense that they would have a large effect on happiness.

# Final Analysis

From looking at this data, we can say that if a country wants to improve their citizens happiness, they should look to raise their gdp, improve their education system and improve their healthcare system. Obviously this is not surprising and is very difficult to do in poor countries.


